---
title: "ant388notes19feb24"
output: html_document
date: "2024-02-19"
---
### For Wednesday
Module 15 (classical null hypothesis significance testing) and Module 13
(Bernoulli binomial, and Poisson distributions)
  Probability mass vs. Probability density functions

### Rmd advice

- Text outside of code blocks
- Formatting on different levels
- R code chunk options
    - echo
    - eval
    - message
    - include
    - fig-height
    - out-width

### In class: Module 14
```{r}
library(tidyverse)
install.packages("manipulate")
library(manipulate)
library(mosaic)
library(radiant)
```

Sampling distributions have pretty well defined statistical properties.

If population variance / standard deviation are known (rare!)...
    ...standard deviation of population/sqrt(sample size)
    ...= expected standard error in a sample

If we can sample population repeatedly to generate a sampling distribution...
    ...standard deviation of sampling distribution

If we have a single sample...
    ...standard deviation of sample/sqrt(sample size)
    ...= an estimate of standard error


```{r}
library(radiant)
n <- 10000 #Written in order to easily change values for n, mu, sd
mu <- 2
sd <- 4
d <- rnorm(n,mu,sd)

mean(d)
sd(d) #sample standard deviation
4/sqrt(1000) #standard deviation of population/sqrt(sample size)
sd(d)/sqrt(1000) #standard deviation of sample/sqrt(sample size)

reps <- 1000
samp_dist <- do(reps) * mean(rnorm(n,mu,sd))
(se_samp_dist <- sd(samp_dist$mean)) #based on sampling distribution
```

### Calculating theoretical confidence intervals

Generating a vector of random numbers and calculating the relevant summary statistics:
```{r}
n <- 10000 #Written in order to easily change values for n, mu, sd
mu <- 25
sd <- 10

set.seed(1)

v <- rnorm(n,mu,sd)

s <- sample(v,size=40,replace=FALSE) 
#Sampling 40 observations from vector without replacement

(m<-mean(s))
```
Assuming the population standard deviation is known..
```{r}
(se <- sd/sqrt(length(s)))
```

Assuming the population standard deviation is unknown...
```{r}
(se <- sd(s)/sqrt(length(s)))
```


```{r}
percent_ci <- 95
alpha <- 1 - percent_ci/100 #alpha = 0.05
lower <- m + qnorm(alpha/2) * se #where the qnorm(alpha/2) is the 2.5% quantile 
                                # of the standard normal distribution
upper <- m + qnorm(1-alpha/2) * se #where the qnorm(1-alpha/2) is the 97.5% 
                                  # quantile of the standard normal distribution
(ci <- c(lower,upper))
```

Psst! There are potentially simpler ways to do this in Module 14...

...work through examples in Module 14!

```{r}

```


