---
title: "ant388notes26feb24"
output: html_document
date: "2024-02-26"
---

```{r}
library(tidyverse)
library(mosaic)
library(readr)
library(dplyr)
library(ggplot2)
```

```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/tbs-2006-2008-ranges.csv"
d <- read_csv(f, col_names = TRUE)

```
Starting exercise:
-Determine mean, SD, SE in "kernel95" HR size for each sex
-Boxplot comparing "kernel95" size by sex
-For each sex, generate a bootstrap distribution, resampling data within each sex 10,000 times
-Plot an appropriate normal distribution over the boostrap distrubution
-Determine 95% CI in home range size for each sex
    -Using quantile() method applied to your bootstrap distribution
     -Using theory-based method based on qnorm()
```{r}
female_data <- subset(d,sex.code==1,kernel95)
male_data <- subset(d,sex.code==2,kernel95)
```

For females
```{r}
(x_f <- female_data$kernel95)
(n_f <- length(female_data$kernel95))
(m_f <- mean(female_data$kernel95))
(sd_f <- sd(female_data$kernel95))
(se_f <- sd(female_data$kernel95)/sqrt(n_f))
```
Generating a bootstrap distribution
```{r}
(x_f <- female_data$kernel95)
(n_f <- length(female_data$kernel95))
(m_f <- mean(female_data$kernel95))
(sd_f <- sd(female_data$kernel95))
(se_f <- sd(female_data$kernel95)/sqrt(n_f))
set.seed(1)

v_f <- rnorm(x_f,m_f,sd_f)
s_f <- sample(v_f,size=n_f,replace=FALSE)

#bootstrapping a CI
n_boot_f <- 10000
boot_f <- vector(length = n_boot) #set up a dummy variable
n_f <- length(s_f)
# the size of each bootstrap sample should be equivalent to sample
for (i in 1:n_boot_f){
  boot_f[[i]] <- mean(sample(s_f,n_f,replace = TRUE))}
hist(boot_f, breaks=25, ylim = c(0,1600), xlab = "Mean", main = "Female Bootstrap Sampling Distribution")
plotDist("norm",318.8,65.7)
```
```{r}
lower_f <- quantile(boot_f,0.025)
upper_f <- quantile(boot_f,0.975)
(ci_boot_f <- c(lower_f,upper_f))
```



For males
```{r}
(x_m <- male_data$kernel95)
(n_m <- length(male_data$kernel95))
(m_m <- mean(male_data$kernel95))
(sd_m <- sd(male_data$kernel95))
(se_m <- sd(male_data$kernel95)/sqrt(n))
```
```{r}
(x_m <- male_data$kernel95)
(n_m <- length(male_data$kernel95))
(m_m <- mean(male_data$kernel95))
(sd_m <- sd(male_data$kernel95))
(se_m <- sd(male_data$kernel95)/sqrt(n))
set.seed(1)

v_m <- rnorm(x_m,m_m,sd_m)
s_m <- sample(v_m,size=n_m,replace=FALSE)

#bootstrapping a CI
n_boot_m <- 10000
boot_m <- vector(length = n_boot_m) #set up a dummy variable
n_m <- length(s_m)
# the size of each bootstrap sample should be equivalent to sample
for (i in 1:n_boot_m){
  boot_m[[i]] <- mean(sample(s_m,n_m,replace = TRUE))}
hist(boot_m, breaks=25, ylim = c(0,1600), xlab = "Mean", main = "Male Bootstrap Sampling Distribution")
plotDist("norm",429.7,58.3)
```
```{r}
lower_m <- quantile(boot_m,0.025)
upper_m <- quantile(boot_m,0.975)
(ci_boot_m <- c(lower_m,upper_m))
```



```{r}
(p <- ggplot(data = d, aes(x = d$sex, y = d$kernel95)) +
  geom_boxplot(na.rm = TRUE) + 
  theme(axis.text.x = element_text(angle = 90)) + #Specifying axis labels
  ylab("kernel95") + 
  xlab("Sex"))
```
```{r}
Sp2 <- ((n_f-1)*sd_f^2)+(n_m-1)*sd_m^2/(n_f+n_m-2)
T = (m_m-m_f-0)/(sqrt(Sp2*(1/n_f+1/n_m)))
t.test(female_data$kernel95,male_data$kernel95, alternative = c("two.sided"))
```
Something is wrong in my Sp2 and/or T calculations by hand above...

**Read Ch 8 and Ch 9 of Ismay and Kim!!!**
**Module 16**
