---
title: "ant388notes4mar24"
output: html_document
date: "2024-03-04"
---

*Ordinary Least Squares*
  Find coefficients that minimize "the residuals", i.e. the difference between
  observed and expected y values.

```{r}
library(manipulate)
library(tidyverse)
install.packages("lmodel2")
install.packages("broom")
library(patchwork)
library(infer)
```

```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/zombies.csv"
d <- read_csv(f,col_names = TRUE)
(head(d))
```

```{r}
d <- mutate(d, centered_height = height - mean(height))
d <- mutate(d, centered_weight = weight - mean(weight))

p1 <- ggplot(data=d, aes(x=weight,y=height)) + geom_point()
p2 <- ggplot(data=d, aes(x=centered_weight,y=centered_height)) + geom_point()

(p1 + p2)
```

Generating a function that plots a scatter of points and a line that runs through that scatter of points which we can determine the equation of:
```{r}
slope.test <- function(beta1, data) {
    g <- ggplot(data = data, aes(x = centered_weight, y = centered_height))
    g <- g + geom_point()
    g <- g + geom_abline(intercept = 0, slope = beta1, size = 1, colour = "blue",
        alpha = 1/2)
    ols <- sum((data$centered_height - beta1 * data$centered_weight)^2)
    g <- g + ggtitle(paste("Slope = ", beta1, "\nSum of Squared Deviations = ", round(ols,
        3)))
    g
}

(manipulate(slope.test(beta1, data = d), beta1 = slider(-1, 1, initial = 0, step = 0.005)))
#able to get interactive slider typing in console with new Rscript doc
```
Calculating β1 by hand...
```{r}
(β1 <- cov(d$weight,d$height)/var(d$weight))
(β0 <- mean(d$height) - β1*(mean(d$weight)))
```

**Challenge**

```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Street_et_al_2017.csv"
d <- read_csv(f,col_names = TRUE)
```

Plot brain size (ECV) as a function of social group size, longevity, juvenile period length, and reproductive lifespan (separate plots for each). 
```{r}
(p1 <- ggplot(data=d, aes(x=Group_size,y=ECV)) + geom_point())
```
```{r}
(p2 <- ggplot(data=d, aes(x=Longevity,y=ECV)) + geom_point())
```
```{r}
(p3 <- ggplot(data=d, aes(x=Weaning,y=ECV)) + geom_point())
```
```{r}
(p4 <- ggplot(data=d, aes(x=Repro_lifespan,y=ECV)) + geom_point())
```

Derive by hand the ordinary least squares regression coefficients B1 and B0 for ECV ~ social group size
```{r}
s <- d |> filter(!is.na(Group_size),!is.na(ECV))
(β1 <- cov(s$Group_size,s$ECV)/var(s$Group_size))
(β0 <- mean(s$ECV) - β1*(mean(s$Group_size)))
```

Confirm that you get the same results using the lm() function
  m <- lm(formula = ECV ~ Group_size, data=d)
  m <- lm(ECV ~ Group_size,d)
```{r}
(m <- lm(ECV ~ Group_size,d))
```
```{r}
(summary(m))
```
```{r}
(names(m))
```
```{r}
(m$coefficients)
```
```{r}
(m$model)
```
```{r}
(broom::tidy(m))
```
```{r}
(broom::glance(m))
```
```{r}
library(broom)
```

Repeat the analysis above for different groups of primates (Catarrhini, Platyrrhini, Strepsirhini) separately. Do you regression coefficients differ? How might you determine this?

For catarrhines...
```{r}
c <- d |> filter(!is.na(Group_size),!is.na(ECV),Taxonomic_group=="Catarrhini")
(β1 <- cov(c$Group_size,c$ECV)/var(c$Group_size))
(β0 <- mean(c$ECV) - β1*(mean(c$Group_size)))
```

For platyrrhines...
```{r}
p <- d |> filter(!is.na(Group_size),!is.na(ECV),Taxonomic_group=="Platyrrhini")
(β1 <- cov(p$Group_size,p$ECV)/var(p$Group_size))
(β0 <- mean(p$ECV) - β1*(mean(p$Group_size)))
```

For strepsirhines...
```{r}
s <- d |> filter(!is.na(Group_size),!is.na(ECV),Taxonomic_group=="Strepsirhini")
(β1 <- cov(s$Group_size,s$ECV)/var(s$Group_size))
(β0 <- mean(s$ECV) - β1*(mean(s$Group_size)))
```
Spend some trying working on a permutation test to evaluate whether the above 
regression coefficients differ...



