---
title: "ant388cnotes7Feb24"
output: html_document
date: "2024-02-07"
---
# Functions and Flow Control

## Module 11
Read and work through examples by Mon. 2/12

Work on finding a paper and dataset for your "Data Analysis Replication" assignment

Submit URL for your repo for Exercise 3 (due Mon. 2/12)

Start looking at Exercise 4 (due Mon. 
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
```

```{r}
d <- read.csv("C:/Users/hijam/Downloads/IMDB-movies.csv")
```
**Conditionals**

if (test) {do this} else {do that}

ifelse(test, od this if true, do this if false)

if_else(test, do this if true, do this if false)

case_when (
case1 ~ result1,
case2 ~ result2,
case3 ~ result3,
TRUE ~ all others
)

```{r}
d <- d|> mutate(comedy = if_else(str_detect(genres,"Comedy"), TRUE, FALSE))

```
str_detect() returns a logical vector with TRUE for each element of string (genres) that matches pattern ("Comedy") and FALSE otherwise

```{r}
s <- d|> filter(comedy == TRUE)
nrow(s)
```
The above technically works.
Using summarise(), I can read the number of movies and the number of comedies.
```{r}
s <- d|> summarise(count = n(), num_comedy = sum(comedy, na.rm = TRUE))
```
```{r}
d <- d|> mutate(ranking = case_when(
  averageRating <= 3.3  ~ "low",
  averageRating <= 6.7 ~ "med",
  averageRating > 6.7 ~ "high",
))

```
```{r}
g <- d|> group_by(ranking)
s <- g|> summarise(count = n(), avg_time = mean(runtimeMinutes, na.rm = TRUE))
```
**Loops**
Another means of flow control

Anything we can apply to a whole dataframe, we can loop through each row in a dataframe

Loops (running things iteratively) can slow down processing

In general, don't do loops in code unless real utility

i means index

for (i in 1:. . . .) {do this}

*For this next challenge, going to loop through all the rows in the dataset and print out the cumulative running time of movies*

First, let's practice writing a loop
```{r}
for(i in 1:nrow(d)){
  print(i)
}
```
### ***Oy gevalt***

```{r}

```
New to resolve an issue below:

t <- 0
for(i in 1:nrow(d)){
  t <- ifelse(is.na(d[i,c("runtimeMinutes")])t, t  
  + d[i,c("runtimeMinutes")])
  print(t)
}

Alternate example:
for (name in c("a","b","c")){
  d <- read_csv(name, colnames = TRUE)
}

```{r}

```

### Walking through Wordle exercise

