---
title: "ant388notes21feb24"
output: html_document
date: "2024-02-21"
---

### Calculating CIs by bootstrapping

n_boot <- 10000
boot <- vector(length=n_boot) #set up a dummy variable to hold our simulations
n <- length(x) #bootstrap sample size will be the same length as our sample data
for (i in 1:n_boot){boot[[i]] <- mean(sample(x,b,replace=TRUE))}

```{r}
library(tidyverse)
library(dplyr)
library(manipulate)
library(mosaic)
x <- 10000
mu <- 25
sigma <- 10
n <- 15
set.seed(1)

v <- rnorm(x,mu,sigma)
s <- sample(v,size=n,replace=FALSE)
(m <- mean(s))
(se <- sd(s)/sqrt(length(s)))

#CI assuming normal distribution
(percent_ci <- 95)
(alpha <- 1 - percent_ci/100)
(lower <- m + qnorm(alpha/2)*se) #lower <- m + qt(p =25,df=14)*se
(upper <- m + qnorm(1-alpha/2)*se)
(ci_norm <- c(lower,upper))

#bootstrapping a CI
n_boot <- 10000
boot <- vector(length = n_boot) #set up a dummy variable
n <- length(s)
# the size of each bootstrap sample should be equivalent to sample
for (i in 1:n_boot){
  boot[[i]] <- mean(sample(s,n,replace = TRUE))}
hist(boot, breaks=25, ylim = c(0,1600), xlab = "Mean", main = "Bootstrap Sampling Distribution")

lower <- quantile(boot,0.025)
upper <- quantile(boot,0.975)
(ci_boot <- c(lower,upper))

```
### Classical hypothesis testing

**Null hypothesis significance testing**
```{r}
library(curl)
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/vervet-weights.csv"
d <- read_csv(f,col_names=TRUE)
summary(d)

mean(d$weight)

mu <- 5
x <- d$weight
n <- length(x)
m <- mean(x)
se <- sd(x)/sqrt(n)

z <- (m-mu)/se
z

p <- 1-pnorm(z)
p

crtiical_val <- qnorm(0.95)
alpha <- 0.05
#one-tailed test
(ci <- m - qnorm(1-alpha, mean=0, sd=1)*se) #lower limit by hand

#does CI include mu? no!

# two-tailed test
(ci <- m + c(-1,1)*qnorm(1-alpha/2, m=0, sd=1)*se) #by hand

#t test
(p <- 1 - pt(z,df=n-1))
(z)

(critical_val <- qt(0.95,df = n-1))
alpha <- 0.05
ci <- c(m-qt(1-alpha, df=n-1)*se,Inf)
ci
(t_stat <- t.test(x=x,mu=mu,alternative="greater"))
```
**Mod15 Challenge1**

```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/woolly-weights.csv"
d <- read_csv(f,col_names=TRUE)
summary(d)
```
```{r}
mu <- 7.2
n <- length(d$weight)
m <- mean(d$weight)
sd <- sd(d$weight)
se <- sd(d$weight)/sqrt(n)
t <- (m-mu)/se

z <- (m-mu)/se
z

p <- 1-pnorm(z)
p

critical_val <- qnorm(0.95)
alpha <- 0.05
```

```{r}
# two-tailed test
(ci <- m + c(-1,1)*qnorm(1-alpha/2, m=0, sd=1)*se) #by hand
```
```{r}
(t_stat <- t.test(d$weight,mu=mu,alternative="two.sided"))
```
```{r}
(p_upper <- 1-pt(abs(t),df=n-1))
(p_lower <- pt(-1*abs(t),df=n-1))
(p <- p_upper + p_lower)
```
**Two sample t test**
Hypothesis test for the difference between two means.

df <- (s2^2/n2 + s1^2/n1)^2/((s2^2/n2)^2/(n2 - 1) + (s1^2/n1)^2/(n1 - 1))
df

**Mod15 Exercise2**

```{r}

```
















