---
title: "ant388notes14feb24"
output: html_document
date: "2024-02-14"
---
```{r}
library(mosaic)
library(radiant)
```

### Big Ideas
**Location**
  - Mean (average)
  - Median
  - Mode
  - Harmonic Mean
  
**Spread**
  Mean Deviation (average of absolute value of deviations from mean)
  Mean Square Deviation (average of squared deviations from mean) = variance
  Standard deviation (square root of variance)

Population Variance = a parameter 
  = SS / N = ∂^2
  N = population size
  sum((x-mean(x))^2)/length(x)
  
Sample Variance = a statistic
  = SS/n - 1 = s^2
  n = sample size
  sum((x-mean(x))^2)/(length(x) - 1)
  var()

Population Standard Deviation = a parameter
  = ∂
  radiant::sdpop()

Sample Standard Deviation = a statistic 
  = s
  sd()
  
```{r}
a <- c(1,2,3,4,5,7,8,9,10)

pop_variance <- function(x){
  sum((x-mean(x))^2)/length(x)
}

pop_variance(a)

sample_variance <- function(x){
  sum((x-mean(x))^2)/(length(x) - 1)
}

sample_variance(a)
```
Some statistics are estimates of the parameters of distributions.

*Let's draw and sample from some mathematical distributions and see how statistics we calculate compare to parameters of distributions*

```{r}
library(mosaic)
mu <- 10
sigma <- 2
plotDist("norm",mean=mu,sd=sigma,xlab="x",ylab="Frequency")
```
```{r}
s <- rnorm(n=10,mean=10,sd=2) #Creating sample that consists of 10 observations drawn out of a normal distribution with a mean of 10 and standard deviation of 2
mean(s)
sd(s)
```
```{r}
s <- rnorm(n=1000,mean=10,sd=2)
mean(s)
sd(s)
```
*Let's create a sampling distribution for the mean of sample drawn from normal distribution with mean of 10 and sd of 2*

```{r}
reps <- 500
sample_dist_mean <-
  do(reps)*mean(rnorm(n=1000,mean=10,sd=2))
str(sample_dist_mean)
#generates a sampling distribution for the mean of our sample
```
```{r}
sample_dist_median <- do(reps)*median(rnorm(n=1000,mean=10,sd=2))
#generates a sampling distribution for the median of our sample
```

**Plots of Sampling Distributions**

```{r}
histogram(~mean,data=sample_dist_mean,xlab="Samp Dist for the Mean")
```
```{R}
histogram(~median,data=sample_dist_median,xlab="Samp Dist for the Median")
```

The mean of a sampling distribution for a particular statistic should be a really good point estimate of the population value for that statistic
```{r}
mean(sample_dist_mean$mean)
```
Standard error (SE) = square root of the variance of the sampling distribution = standard deviation of a sampling distribution
```{r}
se_mean <- sd(sample_dist_mean$mean)
se_median <- sd(sample_dist_median$median)
```
*The above is an example of programming your own **simulation***

  The SE can be used to derive another measure of uncertainty in a statistic's value: the confience interval, or CI
  The CI is this another way of describing a statistic's sampling distribution, and it plays a  central role in basic inferential statistics.
  An interval around our estimate of mean of the sampling distribution for a particular         statistic (usually, mean) and it gives us a range of values into which subsequent estimates of a statistic would be expected to fall some critical proportion of the time, if the sampling      exercise were to be repeated.
  




